---
author: "Therese LaRue"
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
---

# Overview
This file is used to prepare the raw output of GLORIA for downstream analysis. Specifically, this file takes all of the different GLORIA outputs and generates single local, global, and ROI files with descriptor information. 

File inputs (for each GLORIA run, saved in 0-input_files): 

* All data: local, global, and ROI
* Tip data: local 
* Key file (could also be the GLO-Bot run file)

File outputs (saved to 1-working_files):

* Key files
* Merged files: all_roi_data-bind_tidy.csv, all_global_data_bind_tidy.csv, all_local_data_bind_tidy.csv, and tip_local_data_bind_tidy.csv

# Libraries
The following packages were loaded: `library(plyr)`, `library(tidyverse)`, and `library(openxlsx)`.
```{r Libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(plyr)
library(tidyverse)
library(openxlsx)
```

# Import all of the data

## Read in the keys
```{r Data import key, echo=TRUE, message=FALSE, warning=FALSE}
key <- read.csv("0-input_files/GWAS_key.csv", header = TRUE)
head(key)

#ecotype key contains data i curated for my accessions
ecotype_key <- read.xlsx("0-input_files/accessions_data.xlsx", sheet = 1, startRow = 1, colNames = TRUE)
head(ecotype_key)
```

## Read in the entire root data
```{r Data import all, echo=TRUE, message=FALSE, warning=FALSE}
GWA1_all_local <- read.csv("0-input_files/GWA1_all_root-data-local.csv", header = TRUE)
GWA2_all_local <- read.csv("0-input_files/GWA2_all_root-data-local.csv", header = TRUE)
GWA3_all_local <- read.csv("0-input_files/GWA3_all_root-data-local.csv", header = TRUE)
GWA4_all_local <- read.csv("0-input_files/GWA4_all_root-data-local.csv", header = TRUE)
GWA5_all_local <- read.csv("0-input_files/GWA5_all_root-data-local.csv", header = TRUE)
GWA6_all_local <- read.csv("0-input_files/GWA6_all_root-data-local.csv", header = TRUE)
all_local_data_bind <- rbind(GWA1_all_local, GWA2_all_local, GWA3_all_local, GWA4_all_local, GWA5_all_local, GWA6_all_local)
head(all_local_data_bind)

GWA1_all_global <- read.csv("0-input_files/GWA1_all_root-data-global.csv", header = TRUE)
GWA2_all_global <- read.csv("0-input_files/GWA2_all_root-data-global.csv", header = TRUE)
GWA3_all_global <- read.csv("0-input_files/GWA3_all_root-data-global.csv", header = TRUE)
GWA4_all_global <- read.csv("0-input_files/GWA4_all_root-data-global.csv", header = TRUE)
GWA5_all_global <- read.csv("0-input_files/GWA5_all_root-data-global.csv", header = TRUE)
GWA6_all_global <- read.csv("0-input_files/GWA6_all_root-data-global.csv", header = TRUE)
all_global_data_bind <- rbind(GWA1_all_global, GWA2_all_global, GWA3_all_global, GWA4_all_global, GWA5_all_global, GWA6_all_global)
head(all_global_data_bind)

GWA1_all_roi <- read.csv("0-input_files/GWA1_all_root-data-roi.csv", header = TRUE)
GWA2_all_roi <- read.csv("0-input_files/GWA2_all_root-data-roi.csv", header = TRUE)
GWA3_all_roi <- read.csv("0-input_files/GWA3_all_root-data-roi.csv", header = TRUE)
GWA4_all_roi <- read.csv("0-input_files/GWA4_all_root-data-roi.csv", header = TRUE)
GWA5_all_roi <- read.csv("0-input_files/GWA5_all_root-data-roi.csv", header = TRUE)
GWA6_all_roi <- read.csv("0-input_files/GWA6_all_root-data-roi.csv", header = TRUE)
all_roi_data_bind <- rbind(GWA1_all_roi, GWA2_all_roi, GWA3_all_roi, GWA4_all_roi, GWA5_all_roi, GWA6_all_roi)
head(all_roi_data_bind)
```

## Read in the tip data
```{r Data import tips, echo=TRUE, message=FALSE, warning=FALSE}
GWA1_tip <- read.csv("0-input_files/GWA1_tip_root-data-local.csv", header = TRUE)
GWA2_tip <- read.csv("0-input_files/GWA2_tip_root-data-local.csv", header = TRUE)
GWA3_tip <- read.csv("0-input_files/GWA3_tip_root-data-local.csv", header = TRUE)
GWA4_tip <- read.csv("0-input_files/GWA4_tip_root-data-local.csv", header = TRUE)
GWA5_tip <- read.csv("0-input_files/GWA5_tip_root-data-local.csv", header = TRUE)
GWA6_tip <- read.csv("0-input_files/GWA6_tip_root-data-local.csv", header = TRUE)
tip_local_data_bind <- rbind(GWA1_tip, GWA2_tip, GWA3_tip, GWA4_tip, GWA5_tip, GWA6_tip)
head(tip_local_data_bind)
```

## Day numbering
```{r Day numbering, echo=TRUE, message=FALSE, warning=FALSE}
key$accession <- substr(key$label_notes, start = 1, stop = nchar(as.character(key$label_notes))-2)
key$line <- substr(key$label_notes, start = nchar(as.character(key$label_notes)), stop = nchar(as.character(key$label_notes)))

df_dates <- data.frame(experiment_part = character(), start_date = character())
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA1_Part1",start_date = "8/24/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA1_Part2",start_date = "8/25/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA2_Part1",start_date = "9/14/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA2_Part2",start_date = "9/15/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA3_Part1",start_date = "10/04/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA3_Part2",start_date = "10/05/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA4_Part1",start_date = "10/20/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA4_Part2",start_date = "10/21/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA5_Part1",start_date = "11/15/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA5_Part2",start_date = "11/16/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA6_Part1",start_date = "12/12/17 06:00"))
df_dates <- rbind(df_dates, data.frame(experiment_part = "GWA6_Part2",start_date = "12/13/17 06:00"))
df_dates$start_date <- as.POSIXct(df_dates$start_date, format = "%m/%d/%y %H:%M")

df_dates_final <- data.frame(image = all_global_data_bind$image)
df_dates_final$barcode <- substr(as.character(df_dates_final$image), start = 1, stop = 9)
df_dates_final$experiment <- substr(as.character(df_dates_final$image), start = 1, stop = 4)
df_dates_final$rhiz_num <- substr(as.character(df_dates_final$image), start = 6, stop = 9)
df_dates_final$full_date <- as.POSIXct(substr(as.character(df_dates_final$image), start = 11, stop = 25), format = "%Y-%m-%d-%H%M")
df_dates_final$date <- as.POSIXct(substr(as.character(df_dates_final$image), start = 11, stop = 21), format = "%Y-%m-%d")
df_dates_final$plaintime <- substr(as.character(df_dates_final$image), start = 22, stop = 25)
df_dates_final$time <- as.POSIXct(substr(as.character(df_dates_final$image), start = 22, stop = 25), format = "%H%M") 

key$barcode <- as.character(key$barcode)
df_dates_final <- left_join(df_dates_final, key[c("barcode", "label_notes", "other_notes", 
                                                  "experiment_part", "accession", "line")], by = "barcode")
df_dates_final <- left_join(df_dates_final, df_dates, by = "experiment_part")
df_dates_final$hours <- df_dates_final$full_date - df_dates_final$start_date
df_dates_final$approxday <- (df_dates_final$hours/24)+14 #approximate the imaging day which started 14 days after sowing
df_dates_final$day <- as.factor(trunc(df_dates_final$approxday))
df_dates_final$id <- paste(df_dates_final$barcode, df_dates_final$day, sep = "-")
```

# Establish the keys
```{r Establish a key, echo=TRUE, message=FALSE, warning=FALSE}
ecotype_key[ecotype_key=="n.d."] <- NA
head(ecotype_key)

new_key <- df_dates_final %>% 
  select(image, barcode, experiment, rhiz_num, full_date, label_notes, other_notes, 
         experiment_part, accession, line, hours, day, id)
new_key$image_tip <- gsub("invert", "tip", new_key$image)
new_key <- left_join(new_key, ecotype_key %>% select(rhiz_num, ecotype_id_1001, ecotype_id_GWAPP, ecotype_id_busch), by = "rhiz_num")

new_key$ecotypeid <- new_key$ecotype_id_1001
new_key$ecotypeid[is.na(new_key$ecotypeid)] <- new_key$ecotype_id_GWAPP[is.na(new_key$ecotypeid)]
new_key$ecotypeid[is.na(new_key$ecotypeid)] <- new_key$ecotype_id_busch[is.na(new_key$ecotypeid)]

head(new_key)
```

# Add in keys, format, and write files

## Keys
```{r Keys, echo=TRUE, message=FALSE, warning=FALSE}
dir.create("1-working_files")
write.csv(new_key,'1-working_files/new_key.csv', row.names=FALSE)
write.csv(ecotype_key,'1-working_files/ecotype_key.csv', row.names=FALSE)
```

## All ROI
```{r Write all ROI, echo=TRUE, message=FALSE, warning=FALSE}
all_roi_data_bind_tidy <- left_join(all_roi_data_bind, new_key, by = "image")
write.csv(all_roi_data_bind_tidy,'1-working_files/all_roi_data_bind_tidy.csv', row.names=FALSE)
```

## All global
```{r Write all global, echo=TRUE, message=FALSE, warning=FALSE}
all_global_data_bind_tidy <- left_join(all_global_data_bind, new_key, by = "image")
write.csv(all_global_data_bind_tidy,'1-working_files/all_global_data_bind_tidy.csv', row.names=FALSE)
```

## All local
Since GLORIA outputs the angles relative to gravity from 0 to 180, this re-calculates the angles from 0 to 90, where 0 is straight down and 90 is straight out to the side. 
```{r Write all local, echo=TRUE, message=FALSE, warning=FALSE}
all_local_data_bind_tidy <- left_join(all_local_data_bind, new_key, by = "image")
all_local_data_bind_tidy$start_x <- ifelse(all_local_data_bind_tidy$angle < 90, #pointing left
                           (all_local_data_bind_tidy$x + (all_local_data_bind_tidy$length)*sin(abs(all_local_data_bind_tidy$angle-90)*(pi/180))),
                           all_local_data_bind_tidy$x)
all_local_data_bind_tidy$start_y <- all_local_data_bind_tidy$y
all_local_data_bind_tidy$end_x <- ifelse(all_local_data_bind_tidy$angle > 90, #pointing right 
                         (all_local_data_bind_tidy$x + (all_local_data_bind_tidy$length)*(sin(abs(all_local_data_bind_tidy$angle-90)*(pi/180)))), 
                         all_local_data_bind_tidy$x)
all_local_data_bind_tidy$end_y <- all_local_data_bind_tidy$y + abs((all_local_data_bind_tidy$length)*(cos(abs(all_local_data_bind_tidy$angle-90)*(pi/180))))
write.csv(all_local_data_bind_tidy,'1-working_files/all_local_data_bind_tidy.csv', row.names=FALSE)
```

## Tip local
```{r Write tip local, echo=TRUE, message=FALSE, warning=FALSE}
tip_local_data_bind$image <- as.character(tip_local_data_bind$image)
tip_local_data_bind_tidy <- left_join(tip_local_data_bind, new_key, by = c("image"="image_tip"))
tip_local_data_bind_tidy$start_x <- ifelse(tip_local_data_bind_tidy$angle < 90, #pointing left
                           (tip_local_data_bind_tidy$x + (tip_local_data_bind_tidy$length)*sin(abs(tip_local_data_bind_tidy$angle-90)*(pi/180))),
                           tip_local_data_bind_tidy$x)
tip_local_data_bind_tidy$start_y <- tip_local_data_bind_tidy$y
tip_local_data_bind_tidy$end_x <- ifelse(tip_local_data_bind_tidy$angle > 90, #pointing right 
                         (tip_local_data_bind_tidy$x + (tip_local_data_bind_tidy$length)*(sin(abs(tip_local_data_bind_tidy$angle-90)*(pi/180)))), 
                         tip_local_data_bind_tidy$x)
tip_local_data_bind_tidy$end_y <- tip_local_data_bind_tidy$y + abs((tip_local_data_bind_tidy$length)*(cos(abs(tip_local_data_bind_tidy$angle-90)*(pi/180))))
write.csv(tip_local_data_bind_tidy,'1-working_files/tip_local_data_bind_tidy.csv', row.names=FALSE)
```
