---
title: "3-summarize_traits"
author: "Therese LaRue"
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
---

# **Overview**
This file uses the clean files from 2-clean.R to generate a trait matrix. Comparing these methods to other ways to calculate traits can be found in /checking_methods/3-summarize_traits_check.Rmd. 

# **Libraries**
The following packages were loaded: `library(plyr)`, `library(tidyverse)`, `library(geometry)`, `library(plotly)`, `library(ggfortify)` (NOTE: Use R 3.5.2 or later!), `library(egg)`, and `library(corrplot)`.
```{r Libraries, echo=TRUE, message=FALSE, warning=TRUE}
library(plyr)
library(tidyverse)
library(geometry)
```

# **Import files**
```{r Import files, echo=TRUE, message=FALSE, warning=TRUE}
new_key <- read.csv("1-working_files/new_key.csv", header = TRUE)
ecotype_key <- read.csv("1-working_files/ecotype_key.csv", header = TRUE)

clean_traits <- read.csv("1-working_files/clean_traits.csv", header = TRUE)
clean_true_root <- read.csv("1-working_files/clean_true_root.csv", header = TRUE)
clean_ROIs <- read.csv("1-working_files/clean_ROIs.csv", header = TRUE)
```

# **Generate new files**
```{r Generate files, echo=TRUE, message=FALSE, warning=TRUE}
calculated_traits <- unique(new_key %>% 
  select(id, barcode, experiment, 
         rhiz_num, full_date, experiment_part, 
         accession, ecotypeid, line, hours, day))

calculated_traits$rhiz_num <- as.character(calculated_traits$rhiz_num)
calculated_traits$id <- as.character(calculated_traits$id)

ecotype_key$rhiz_num <- as.character(ecotype_key$rhiz_num)
calculated_traits <- left_join(calculated_traits, ecotype_key[,c("rhiz_num",
                                                                 "lat",
                                                                 "long",
                                                                 "swe_or_not")])
```

# **Add in traits calculated while cleaning**
```{r Add in clean traits, echo=TRUE, message=FALSE, warning=TRUE}
clean_traits$id <- as.character(clean_traits$id)
calculated_traits <- left_join(calculated_traits, clean_traits[,c("id",
                                                                  "convexhull", 
                                                                  "xmin",
                                                                  "xmax",
                                                                  "colMeans.x", 
                                                                  "colMeans.y",
                                                                  "depth",
                                                                  "total_length",
                                                                  "weight_avg_angle")], by = "id")
```

# **Calculate additional traits**
```{r Calculate traits, echo=TRUE, message=FALSE, warning=TRUE}
calculated_traits$width <- calculated_traits$xmax-calculated_traits$xmin
calculated_traits$depth_width <- calculated_traits$depth/calculated_traits$width
calculated_traits$density <- calculated_traits$total_length/calculated_traits$convexhull
```

## By image
Depth, length of that image, average angle in that image, and the number of segments in an image. 
```{r By image, echo=TRUE, message=FALSE, warning=TRUE}
clean_true_root$adjusted_angle <- abs(clean_true_root$angle-90)

clean_true_root <- clean_true_root %>% 
  group_by(id) %>% 
  mutate(true_root_tip_depth = max(end_y),
         length_perimg = sum(length),
         weight_avg_angle_perimg = sum(adjusted_angle*length)/sum(length),
         n_segments = length(length))

clean_true_root$id <- as.character(clean_true_root$id)
calculated_traits <- left_join(calculated_traits, unique(clean_true_root[,c("id", 
                                                                            "true_root_tip_depth",
                                                                            "length_perimg",
                                                                            "weight_avg_angle_perimg",
                                                                            "n_segments")]), by = "id")
```

## By barcode
Total length, time, growth amount, rate percent, average growth rate (cm per hour)
```{r By barcode, echo=TRUE, message=FALSE, warning=TRUE}
calculated_traits <- calculated_traits %>% 
  arrange(day) %>% 
  group_by(barcode) %>% 
  mutate(sum_total_length = cumsum(length_perimg),
         diff_hours = hours-lag(hours, default = 0),
         diff_depth = depth-lag(depth, default = 0),
         primary_growthrate_perday = diff_depth/diff_hours,
         primary_growthrate_avg = mean(primary_growthrate_perday, na.rm = TRUE),
         primary_rate_percent = (diff_depth/diff_hours)/depth *100,
         primary_rate_percent_avg = mean(primary_rate_percent, na.rm = TRUE),
         diff_total_length = total_length-lag(total_length, default = 0),
         total_growthrate_perday = diff_total_length/diff_hours,
         total_growthrate_avg = mean(total_growthrate_perday, na.rm = TRUE),
         total_rate_percent = (diff_total_length/diff_hours)/total_length*100)
```

## From ROI

Confirm clean_ROI matches the computed_traits 
```{r By ROI, echo=TRUE, message=FALSE, warning=TRUE}
clean_ROIs$barcode <- substr(clean_ROIs$id, start = 1, stop = 9)
clean_ROIs <- clean_ROIs %>%
  group_by(id) %>%
  mutate(ROI_depth = max(y),
         ROI_xmin = min(x),
         ROI_xmax = max(x),
         ROI_width = ROI_xmax - ROI_xmin)

clean_ROIs$id <- as.character(clean_ROIs$id)
calculated_traits <- left_join(calculated_traits, unique(clean_ROIs[,c("id",
                                                                       "ROI_depth",
                                                                       "ROI_xmin",
                                                                       "ROI_xmax",
                                                                       "ROI_width")]), by = "id")
```

# **Save trait matrix**
```{r save, echo=TRUE, message=FALSE, warning=FALSE}
dir.create("3-analysis")
write.csv(calculated_traits, "3-analysis/calculated_traits.csv")
```
